Real-space computation of $E$/$B$-mode maps I: Formalism, Compact Kernels, and Polarized Filaments

### 2018-12-06 articles/jcap_referee_response/Response to referee/response to referee.tex

\item[{Referee comment: }] The authors derived the radiation and convolution kernels that are shown in Figs. 3 and 4. It is still unclear to me what the E/B power spectra look like from this real-space computation. For example, the authors can make a quick simulation on a small patch or the full-sky to easily validate that the E/B power spectra are correctly recovered from this real space computation.

\item[{Referee comment: }]  Our main aim in this paper was to layout the formalism for real space evaluations of $E/B$ fields from the measured Stokes $Q/U$ parameters and highlight and discuss in detail the two different interpretations, namely the ``radiation" and ``convolution" forms of the real space kernels. Please see the supplementary section at the end of this response, where we present preliminary results from real space evaluation of E/B maps and the corresponding power spectra.\\

\item[{Referee comment:}] What is the time complexity of the real space transformation? Eq. (3.26) seems to indicate it is actually $O(N_{\rm pix}^2 )$ which is not as efficient as the conventional approaches. If possible, the authors could briefly discuss it.

\item[{Authors response:}] Yes the real space implementation which is perfectly equivalent to conventional methods would scale as $O(N_{\rm pix}^2 )$. However we have noted that the computation time can be significantly improved by reconstructing a filtered versions $E/B$ fields which would scale as $O(\epsilon N_{\rm pix}^2 )$, where $\epsilon$ is a very small number and is determined by the radial cutoff. We are in process of quantifying this relative efficiency and plan to report our finding in a subsequent publication.

@@

\item[{Referee comment: }] Figs 6 and 7 are not legible.
\item[{Authors response: }] The label sizes have been revised to make them more legible.

\end{itemize}

\newpage
\section*{Supplementary}
We note that, many of the questions raised by the referee, are relevant, but beyond the scope of the work presented in the article. Many of these questions form the topic of our follow up work, where we discuss the implementation details of the real space operator, their comparison to conventional approaches, E/B mixing on masked skies using these real space kernels, etc.

\textit{In this supplementary section we present some preliminary results which basically demonstrates that the real space E/B separation works.}

We evaluate the $E/B$ maps from simulated Stokes Q/U maps using the real space kernels and finally evaluate the corresponding power spectra. 
%
\begin{figure}[h]
\centering
\subfigure[\label{fig:fbeta}]{\includegraphics[width=0.45\columnwidth]{modified_radial_kernel.pdf}}
\subfigure[\label{fig:gl}]{\includegraphics[width=0.45\columnwidth]{effective_beam.pdf}}
\caption{}
%\label{fig:rad_ker_decay}
\end{figure}
%
For this exercise, we work at a Healpix resolution of ${\textrm Nside}=256$. We simulate CMB polarization map using power spectra computed with CAMB with fiducial cosmological parameters and B-modes corresponding to $r=0.001$. We evaluate the $E/B$ maps using real space kernels on a patch of $10 ^{\circ} \times 25^{\circ}$, centered at zero galactic latitude and the long axis of the patch aligned with the equator. 
%
\begin{figure}[h]
\centering
\subfigure[Real space B]{\includegraphics[width=0.45\columnwidth]{real_space_B.pdf}}
\subfigure[True B]{\includegraphics[width=0.45\columnwidth]{true_B.pdf}}
\caption{The figure on the left depicts the B-mode maps evaluated using the modified real space kernel while that on the right depicts the true B-mode map.}
\label{fig:compare_B}
\end{figure}
%
We evaluate the E/B maps using the real space operator, in which we modify the radial kernel by imposing a radial cut-off of $\sim 22^{\circ}$.  The radial kernel used for constructing the real space operator is shown in Fig. \ref{fig:fbeta} while the effective harmonic space filter corresponding to this radial filter is depicted in Fig.~\ref{fig:gl}.

The resultant $B$ mode maps evaluated using the real space method and true B-mode map evaluated using the conventional method is shown in Fig~\ref{fig:compare_B}. 
%
\begin{figure}[!h]
\centering
\subfigure[E-mode power spectra]{\includegraphics[width=0.45\columnwidth]{real_space_ee_spectra.pdf}}
\subfigure[B-mode power spectra]{\includegraphics[width=0.45\columnwidth]{real_space_bb_spectra.pdf}}
\caption{In these plots, the black line depicts the theory power spectrum, the blue line depict the power spectrum evaluated from the true E/B maps while the red dashed lines depict the power spectrum evaluted from real space evaluated E/B maps.}
\label{fig:compare_spectra}
\end{figure}
%
Finally the power spectra estimated from $E/B$ maps evaluated using these real space operators and their comparison to the true power spectrum is depicted in Fig.~\ref{fig:compare_spectra}.

Note that in the results presented here, we compute the $E/B$ maps on the square patch for which we have used Stokes parameters outside the mask boundary. Hence via the results presented here, we do not claim to address any issues relating to masking induced E/B mixing.

### 2018-12-06 articles/jcap_referee_response/modified article jcap/article.tex
refinements to referee report and revised draft

Near the edges of the domain $\beta \rightarrow 0,\pi$ the window function partly falls outside the domain and this is in part responsible for the scaling deviating from the power law, but it is also important to note that ${}_{\mm}f$ and ${}_{\md}f$ need to vanish at $\beta=0$ and therefore the sliding average in the vicinity of these points cannot have the same power law behavior.
%---------------------------------------------------------------------------------------------------------
%\include{generalizedeb}
\section{Generalized polarization operators and recovering standard power spectra}\label{sec:generalized_operators}  
With our better understanding of the radial part of the kernel for CMB polarization, we can write down generalized $E/B$-like fields that depend on a different radial function, even one that we specify to have compact support.
The spin symmetry constrains the the azimuthal part of the real space kernels to be of the form $\sim e^{\pm i2 \alpha}$.  The radial parts of the standard operators are determined by the sum over spherical harmonics and varies as a function of the band limit. It is here that we may potentially choose alternate forms for the radial functions to suit certain kind of analysis.

>>>>

Near the edges of the domain $\beta \rightarrow 0,\pi$ the window function partly falls outside the domain and this is in part responsible for the scaling deviating from the power law, but it is also important to note that ${}_{\mm}f$ and ${}_{\md}f$ need to vanish at $\beta=0$ and therefore the sliding average in the vicinity of these points cannot have the same power law behavior.
%---------------------------------------------------------------------------------------------------------
%\include{generalizedeb}
\section{Generalized polarization operators and recovering standard power spectra}\label{sec:generalized_operators}  
\addtext{In this section we relate our new found knowledge about the radial part of the kernel to the standard CMB polarization angular power spectra.} With our better understanding of the radial part of the kernel for CMB polarization, we can write down generalized $E/B$-like fields that depend on a different radial function, even one that we specify to have compact support.
The spin symmetry constrains the the azimuthal part of the real space kernels to be of the form $\sim e^{\pm i2 \alpha}$.  The radial parts of the standard operators are determined by the sum over spherical harmonics and varies as a function of the band limit. It is here that we may potentially choose alternate forms for the radial functions to suit certain kind of analysis.
